---
title: "Lineup Analysis"
format: html
editor: source
embed-resources: true
code-fold: true
execute: 
  echo: true
  message: false
  warning: false
---


```{r setup}
#| include: false
library(tidyverse)
library(cowplot)
library(digest)

library(readr)

library(lme4)
library(emmeans)
```

# Data

```{r, eval = T}
lineup_model_data <- read_csv("lineup-model-data.csv") |> 
  mutate(conf_level = factor(conf_level, levels = c("Very Certain", "Certain", "Neutral", "Uncertain", "Very Uncertain"))) |> 
  mutate(conf_level_num = as.numeric(conf_level))
```

```{r}
lineup_model_data |> 
  select(participant_id, age, gender, academic_study, dataset_id, curvature, scale, correct, response_no, binary_response, conf_level, choice_reason) |> 
  head() |> 
  knitr::kable()
```

# Visual of Observed Results

+ There are `r length(unique(lineup_model_data$participant_id)) %>% as.numeric` participants and `r nrow(lineup_model_data) %>% as.numeric()` lineups complete. 
<!-- + Each plot was evaluated between ` min(ploteval_summary$count) %>% as.numeric() %>% round(2)` and ` max(ploteval_summary$count)  %>% as.numeric()  %>% round(2)` times (Mean: ` mean(ploteval_summary$count)  %>% as.numeric()  %>% round(2)`, SD: ` sd(ploteval_summary$count)  %>% as.numeric()  %>% round(2)`). -->
+ Participants correctly identified the target panel in `r round(as.numeric(mean(lineup_model_data$binary_response[lineup_model_data$scale == 'linear']))*100, 2)`% of the `r nrow(lineup_model_data[lineup_model_data$scale == 'linear',]) %>% as.numeric()` lineup evaulations made on the linear scale and `r round(as.numeric(mean(lineup_model_data$binary_response[lineup_model_data$scale == 'log']))*100, 2)`% of the `r nrow(lineup_model_data[lineup_model_data$scale == 'log',]) %>% as.numeric()` lineup evaluations made on the log scale.

```{r}
lineup_model_data |> 
  mutate(target = factor(target, levels = c("H", "M", "E")),
         null = factor(null, levels = c("H", "M", "E"))) |> 
  ggplot(aes(x = scale,
             fill = as.factor(binary_response))) +
  geom_bar(stat = "count",
           position = "fill") +
  facet_grid(null ~ 
             target,
             labeller = label_both) +
  scale_fill_brewer(palette = "Dark2", direction = -1, label = c("Inorrect", "Correct")) +
  scale_y_continuous(label = scales::percent) +
  labs(y = "",
       x = "",
       fill = "Response") +
  theme_test()
```

# Analysis

Target plot identification was analyzed using the `glmer` function in the `lme4` R package. 
Estimates and odds ratio comparisons between the log and linear scales were calculated using the `emmeans` R package.
Each lineup plot evaluated was assigned a value based on the participant response (correct = 1, not correct = 0).
Define $Y_{ijkl}$ to be the event that participant $l = 1,...,N_{participant}$ correctly identifies the target plot for data set $k = 1,2$ with curvature combination $j = 1,2,3,4,5,6$ plotted on scale $i = 1,2$.
The binary response was analyzed using a generalized linear mixed model following a binomial distribution with a logit link function with a row-column blocking design accounting for the variation due to participant and data set respectively as 
\begin{equation}
\text{logit }P(Y_{ijk}) = \eta + \delta_i + \gamma_j + \delta \gamma_{ij} + s_l + d_k
\end{equation}
\noindent where

+ $\eta$ is the baseline average probability of selecting the target plot
+ $\delta_i$ is the effect of scale $i = 1,2$
+ $\gamma_j$ is the effect of curvature combination $j = 1,2,3,4,5,6$
+ $\delta\gamma_{ij}$ is the two-way interaction between the $i^{th}$ scale and $j^{th}$ curvature combination
+ $s_l \sim N(0,\sigma^2_\text{participant})$is the random effect for participant characteristics
+ $d_k \sim N(0,\sigma^2_{\text{data}})$ is the random effect for data specific characteristics. 

\noindent We assume that random effects for data set and participant are independent.

```{r}
glmm_mod <- glmer(binary_response ~ curvature*scale + 
                            (1 | participant_id) +
                            (1 | dataset_id),
                  data = lineup_model_data,
                  family = binomial(link = "logit"))

lineup_anova <- car::Anova(glmm_mod) |> 
  broom::tidy()
# summary(glmm_mod)

lineup_anova |> 
  knitr::kable(digits = 3, caption = "Lineups GLMM ANOVA")
```

```{r, eval = F, include = F, fig.align = 'center', fig.width = 12, fig.height = 6}
lineup_lsmeans <- emmeans(glmm_mod, specs = ~ "curvature:scale", type = "response") |> 
  as_tibble()

lineup_lsmeans |> 
  mutate(curvature = factor(curvature, levels = c("t-E_n-H", "t-H_n-E", "t-E_n-M", "t-M_n-E", "t-M_n-H", "t-H_n-M"))) %>%
  ggplot(aes(x = curvature, y = prob, group = scale)) +
  geom_bar(stat = "identity", fill = "lightgray") +
  # geom_text(aes(y = asymp.UCL + 0.05)) +
  facet_wrap(~scale, ncol = 1) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.1) +
  theme_bw() +
  scale_y_continuous("Probability of target panel detected", limit = c(-0.05,1.1), breaks = seq(0,1,0.2)) +
  scale_x_discrete("Curvature")
```


```{r, fig.align = 'center', fig.width = 12, fig.height = 6}
lineups_or <- emmeans(glmm_mod, ~ scale | curvature, type = "response") |> 
  pairs(reverse = T, infer = c(T,T)) |> 
  as_tibble()

# write.csv(lineups_or, here("analyses/01-lineups/results/lineups-odds-ratios.csv"), row.names = F, na =)

dodge <- position_dodge(width=0.9)
lineups_or_plot <- lineups_or %>%
  separate(curvature, into = c(NA, "target", NA, "null")) %>%
  mutate(null = factor(null, levels = c("E", "M", "H"), labels = c("High Curvature", "Medium Curvature", "Low Curvature")),
         target = factor(target, levels = c("E", "M", "H"), labels = c("High Curvature", "Medium Curvature", "Low Curvature"))) %>%
  ggplot(aes(x = odds.ratio, y = null, color = target, shape = target)) + 
  geom_point(position = dodge, size = 3) + 
  geom_errorbar(aes(xmin = asymp.LCL, xmax = asymp.UCL), position = dodge, width = .1) +
  geom_vline(xintercept = 1) +
  theme_bw()  +
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        legend.key.size = unit(1, "line"),
        legend.position = "bottom"
  ) +
  scale_y_discrete("Null Panel Difficulty", position = "left") +
  scale_x_continuous("Odds ratio (on log scale) \n (Log vs Linear)", trans = "log10", labels = scales::comma) + 
  scale_color_manual("Target Panel Difficulty", values = c("#004400", "#116611", "#55aa55")) + 
  scale_shape_discrete("Target Panel Difficulty")

lineups_or_plot
```

# Confidence Level Analysis

```{r}
#| fig-width: 12
#| fig-height: 8
# lineup_model_data |> 
#   select(participant_id, dataset_id, curvature, scale, binary_response, conf_level, conf_level_num) |> 
#   head() |> 
#   knitr::kable()

# levels(lineup_model_data$conf_level)

lineup_model_data |> 
  mutate(target = factor(target, levels = c("H", "M", "E")),
         null = factor(null, levels = c("H", "M", "E"))) |> 
  ggplot(aes(x = as.factor(binary_response),
             fill = conf_level)) +
  geom_bar(stat = "count",
           position = "fill",
           color = "black") +
  facet_grid(null ~ 
             scale + target,
             labeller = label_both) +
  scale_fill_brewer(palette = "PuOr", direction = -1) +
  scale_y_continuous(label = scales::percent) +
  scale_x_discrete(label = c("Incorrect", "Correct")) +
  labs(y = "",
       x = "",
       fill = "Confidence Level") +
  theme_test()
```

Assuming Normality...

```{r}
# normal
conf_level_num_mod <- lmer(conf_level_num ~ curvature*scale*binary_response + 
                            (1 | participant_id) +
                            (1 | dataset_id),
                  data = lineup_model_data)

car::Anova(conf_level_num_mod) |> 
  broom::tidy()
# summary(conf_level_num_mod)

emmeans(conf_level_num_mod, specs = ~ binary_response | curvature:scale) |> 
  pairs() |> 
  as_tibble() |> 
  mutate(p.value = scales::pvalue(p.value))

emmeans(conf_level_num_mod, specs = ~ scale | curvature:binary_response) |> 
  pairs() |> 
  as_tibble() |> 
  mutate(p.value = scales::pvalue(p.value))

emmeans(conf_level_num_mod, specs = ~ scale | binary_response) |> 
  pairs() |> 
  as_tibble() |> 
  mutate(p.value = scales::pvalue(p.value))
```

assuming multinomial...

```{r}
# nnet
conf_level_mod <- nnet::multinom(conf_level ~ curvature*scale*binary_response, #+ 
                            # (1 | participant_id) +
                            # (1 | dataset_id),
                  data = lineup_model_data,
                  trace = F)

car::Anova(conf_level_mod) |> 
  broom::tidy()
# summary(conf_level_mod)

# emmeans(conf_level_mod, specs = ~ binary_response | curvature:scale, type = "response") |> 
#   pairs() |> 
#   as_tibble() |> 
#   mutate(p.value = scales::pvalue(p.value))
# 
# emmeans(conf_level_mod, specs = ~ scale | curvature:binary_response, type = "response") |> 
#   pairs() |> 
#   as_tibble() |> 
#   mutate(p.value = scales::pvalue(p.value))
# 
# emmeans(conf_level_mod, specs = ~ scale | binary_response, type = "response") |> 
#   pairs() |> 
#   as_tibble() |> 
#   mutate(p.value = scales::pvalue(p.value))

# mlogit
# conf_level_mod2 <- mlogit::mlogit(conf_level ~ curvature*scale + 
#                             (1 | participant_id) +
#                             (1 | dataset_id),
#                   data = lineup_model_data)
# 
# car::Anova(conf_level_mod2) |> 
#   broom::tidy()
# summary(conf_level_mod2)
```
