---
title: "Lineup Analysis"
format: html
editor: source
embed-resources: true
code-fold: true
execute: 
  echo: true
  message: false
  warning: false
---

```{r setup}
#| include: false
library(tidyverse)
library(cowplot)
library(digest)
library(ggmosaic)
library(ggh4x)
library(gridExtra)

library(readr)

library(lme4)
library(emmeans)

# install.packages("devtools")
# devtools::install_github("heike/vinference")
library(vinference)
```

# Data

```{r}
lineup_model_data <- read_csv("lineup-model-data.csv") |> 
  mutate(conf_level = factor(conf_level, levels = c("Very Certain", "Certain", "Neutral", "Uncertain", "Very Uncertain"))) |> 
  mutate(conf_level_num = as.numeric(conf_level))
```

```{r}
lineup_model_data |> 
  select(participant_id, age, gender, academic_study, dataset_id, curvature, scale, correct, response_no, binary_response, conf_level, choice_reason) |> 
  head() |> 
  knitr::kable()
```

# Visual of Observed Results

+ There are `r length(unique(lineup_model_data$participant_id)) %>% as.numeric` participants and `r nrow(lineup_model_data) %>% as.numeric()` lineups complete. 

+ Participants correctly identified the target panel in `r round(as.numeric(mean(lineup_model_data$binary_response[lineup_model_data$scale == 'linear']))*100, 2)`% of the `r nrow(lineup_model_data[lineup_model_data$scale == 'linear',]) %>% as.numeric()` lineup evaulations made on the linear scale and `r round(as.numeric(mean(lineup_model_data$binary_response[lineup_model_data$scale == 'log']))*100, 2)`% of the `r nrow(lineup_model_data[lineup_model_data$scale == 'log',]) %>% as.numeric()` lineup evaluations made on the log scale.

```{r, eval = T}
lineup_model_data <- read_csv("lineup-model-data.csv") |> 
  mutate(conf_level = factor(conf_level, levels = c("Very Certain", "Certain", "Neutral", "Uncertain", "Very Uncertain"))) |> 
  mutate(conf_level_num = as.numeric(conf_level))
```

```{r}
simData <- read.csv("../data/difficulty-comparison-data.csv")

tE_nM <- simData  |>
  mutate(Target = "High",
         Null = "Medium") |>
  filter(curvature %in% c("E", "M")) |>
  mutate(curvature = factor(curvature, levels = c("E", "M", "H"))) |>
  ggplot(aes(x = x, y = y, color = curvature, linetype = curvature)) +
  geom_line(linewidth = 1) +
  theme_test() +
  theme(aspect.ratio = 1) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none") +
  scale_x_continuous(limits = c(0,20)) +
  scale_color_manual(values = c("green3", "black")) +
  scale_linetype_manual(values = c("solid", "dashed"))

tE_nH <- simData  |>
  mutate(Target = "High",
         Null = "Low") |>
  filter(curvature %in% c("E", "H")) |>
  mutate(curvature = factor(curvature, levels = c("E", "M", "H"))) |>
  ggplot(aes(x = x, y = y, color = curvature, linetype = curvature)) +
  geom_line(linewidth = 1) +
  theme_test() +
  theme(aspect.ratio = 1) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none") +
  scale_x_continuous(limits = c(0,20)) +
  scale_color_manual(values = c("green3", "black")) +
  scale_linetype_manual(values = c("solid", "dashed"))

tM_nE <- simData  |>
  mutate(Target = "Medium",
         Null = "High") |>
  filter(curvature %in% c("E", "M")) |>
  mutate(curvature = factor(curvature, levels = c("E", "M", "H"))) |>
  ggplot(aes(x = x, y = y, color = curvature, linetype = curvature)) +
  geom_line(linewidth = 1) +
  theme_test() +
  theme(aspect.ratio = 1) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none") +
  scale_x_continuous(limits = c(0,20)) +
  scale_color_manual(values = c("black", "green3")) +
  scale_linetype_manual(values = c("dashed", "solid"))

tM_nH <- simData  |>
  mutate(Target = "Medium",
         Null = "Low") |>
  filter(curvature %in% c("H", "M")) |>
  mutate(curvature = factor(curvature, levels = c("E", "M", "H"))) |>
  ggplot(aes(x = x, y = y, color = curvature, linetype = curvature)) +
  geom_line(linewidth = 1) +
  theme_test() +
  theme(aspect.ratio = 1) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none") +
  scale_x_continuous(limits = c(0,20)) +
  scale_color_manual(values = c("green3", "black")) +
  scale_linetype_manual(values = c("solid", "dashed"))

tH_nE <- simData  |>
  mutate(Target = "Low",
         Null = "High") |>
  filter(curvature %in% c("E", "H")) |>
  mutate(curvature = factor(curvature, levels = c("E", "M", "H"))) |>
  ggplot(aes(x = x, y = y, color = curvature, linetype = curvature)) +
  geom_line(linewidth = 1) +
  theme_test() +
  theme(aspect.ratio = 1) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none") +
  scale_x_continuous(limits = c(0,20)) +
  scale_color_manual(values = c("black", "green3")) +
  scale_linetype_manual(values = c("dashed", "solid"))

tH_nM <- simData  |>
  mutate(Target = "Low",
         Null = "Medium") |>
  filter(curvature %in% c("H", "M")) |>
  mutate(curvature = factor(curvature, levels = c("E", "M", "H"))) |>
  ggplot(aes(x = x, y = y, color = curvature, linetype = curvature)) +
  geom_line(linewidth = 1) +
  theme_test() +
  theme(aspect.ratio = 1) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none") +
  scale_x_continuous(limits = c(0,20)) +
  scale_color_manual(values = c("black", "green3")) +
  scale_linetype_manual(values = c("dashed", "solid"))

# linear scale

tE_nH_linear <- tE_nH +
  scale_y_continuous(limits = c(10,100))

tE_nM_linear <- tE_nM +
  scale_y_continuous(limits = c(10,100))

tM_nH_linear <- tM_nH +
  scale_y_continuous(limits = c(10,100))

tM_nE_linear <- tM_nE +
  scale_y_continuous(limits = c(10,100))

tH_nM_linear <- tH_nM +
  scale_y_continuous(limits = c(10,100))

tH_nE_linear <- tH_nE +
  scale_y_continuous(limits = c(10,100))

# log scale

library(scales)

tE_nH_log <- tE_nH +
  scale_y_continuous(limits = c(10,100), trans = "log10")

tE_nM_log <- tE_nM +
  scale_y_continuous(limits = c(10,100), trans = "log10")

tM_nH_log <- tM_nH +
  scale_y_continuous(limits = c(10,100), trans = "log10")

tM_nE_log <- tM_nE +
  scale_y_continuous(limits = c(10,100), trans = "log10")

tH_nM_log <- tH_nM +
  scale_y_continuous(limits = c(10,100), trans = "log10")

tH_nE_log <- tH_nE +
  scale_y_continuous(limits = c(10,100), trans = "log10")
```

```{r}
#| fig-width: 8
#| fig-height: 4
scale_level <- c("linear" = "Linear Scale",
                 "log" = "Log Scale")

curvature_diff_level <- c("A - Large Diff" = "Large Curvature Difference",
                          "B - Small Diff" = "Small Curvature Difference",
                          "C - Small Diff" = "Small Curvature Difference"
                          )

curvature_order <- c("t-E_n-H",
                     "t-H_n-E",
                     "t-M_n-H",
                     "t-H_n-M",
                     "t-E_n-M",
                     "t-M_n-E"
                     )

accuracy_plot_linear <- lineup_model_data |> 
  filter(scale == "linear") |> 
  mutate(target = factor(target, levels = c("H", "M", "E")),
         null = factor(null, levels = c("H", "M", "E")),
         curvature_diff = case_when(curvature %in% c("t-E_n-H", "t-H_n-E") ~ "A - Large Diff",
                                    curvature %in% c("t-M_n-H", "t-H_n-M") ~ "B - Small Diff",
                                    curvature %in% c("t-E_n-M", "t-M_n-E") ~ "C - Small Diff"),
         curvature = factor(curvature, levels = curvature_order, 
                            labels = c("Target: High\nNull: Low",
                              "Target: Low \nNull: High",
                              "Target: Medium \nNull: Low",
                              "Target: Low \nNull: Medium",
                              "Target: High \nNull: Medium",
                              "Target: Medium \nNull: High")
                            )
         ) |> 
  ggplot(aes(x = curvature,
             fill = as.factor(binary_response))) +
  geom_bar(stat = "count",
           position = "fill") +
  scale_fill_manual(values = c("darkgray", "#008b00"), label = c("Incorrect", "Correct")) +
  scale_y_continuous(label = scales::percent) +
  scale_x_discrete(position = "top") +
  facet_grid(scale ~ curvature_diff,
             labeller = labeller(scale = scale_level,
                                 curvature_diff = curvature_diff_level),
             scales = "free_x"
             ) +
  labs(y = "",
       x = "",
       fill = "Target Panel \nIdentification",
       title = str_c("Probability of Participants who ", "<span style = 'color:#008b00'>**correctly**</span>", " and ", 
       "<span style = 'color:#808080'>**incorrectly**</span>", " identified the target panel")) +
  theme_test() +
  theme(axis.ticks.x = element_blank(),
        plot.title = ggtext::element_markdown(size = 14),
        legend.position = "none",
        aspect.ratio = 1.5)

accuracy_plot_linear

ggsave(plot = accuracy_plot_linear, filename = "../images/1-linear.svg", dpi = 300, width = 8, height = 5)

```

```{r}
#| fig-width: 8
#| fig-height: 2
# output plots with gridExtra
thumbnails_linear <- grid.arrange(
  tE_nH_linear,
  tH_nE_linear,
  tM_nH_linear,
  tH_nM_linear,
  tE_nM_linear,
  tM_nE_linear,
  ncol = 6
)
thumbnails_linear

ggsave(plot = thumbnails_linear, filename = "../images/1-linear-thumbs.svg", dpi = 300, width = 8, height = 5)
```

```{r}
#| fig-width: 8
#| fig-height: 2
thumbnails_log <- grid.arrange(
  tE_nH_log,
  tH_nE_log,
  tM_nH_log,
  tH_nM_log,
  tE_nM_log,
  tM_nE_log,
  ncol = 6
)

ggsave(plot = thumbnails_log, filename = "../images/1-log-thumbs.svg", dpi = 300, width = 8, height = 5)
```

```{r}
#| fig-width: 8
#| fig-height: 4
accuracy_plot_log <- lineup_model_data |> 
  filter(scale == "log") |> 
  mutate(target = factor(target, levels = c("H", "M", "E")),
         null = factor(null, levels = c("H", "M", "E")),
         curvature_diff = case_when(curvature %in% c("t-E_n-H", "t-H_n-E") ~ "A - Large Diff",
                                    curvature %in% c("t-M_n-H", "t-H_n-M") ~ "B - Small Diff",
                                    curvature %in% c("t-E_n-M", "t-M_n-E") ~ "C - Small Diff"),
         curvature = factor(curvature, levels = curvature_order, 
                            labels = c("Target: High\nNull: Low",
                              "Target: Low \nNull: High",
                              "Target: Medium \nNull: Low",
                              "Target: Low \nNull: Medium",
                              "Target: High \nNull: Medium",
                              "Target: Medium \nNull: High")
                            )
         ) |> 
  ggplot(aes(x = curvature,
             fill = as.factor(binary_response))) +
  geom_bar(stat = "count",
           position = "fill") +
  facet_grid(scale ~ curvature_diff,
             labeller = labeller(scale = scale_level,
                                 curvature_diff = curvature_diff_level),
             scales = "free_x"
             ) +
  scale_fill_manual(values = c("darkgray", "#008b00"), label = c("Incorrect", "Correct")) +
  scale_y_continuous(label = scales::percent) +
  scale_x_discrete(labels = c("Target: High\nNull: Low",
                              "Target: Low \nNull: High",
                              "Target: Medium \nNull: Low",
                              "Target: Low \nNull: Medium",
                              "Target: High \nNull: Medium",
                              "Target: Medium \nNull: High")) +
  labs(y = "",
       x = "",
       fill = "Target Panel \nIdentification") +
  theme_test() +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none",
        aspect.ratio = 1.5)

accuracy_plot_log

ggsave(plot = accuracy_plot_log, filename = "../images/1-log.svg", dpi = 300, width = 8, height = 5)
```

# Analysis

Target plot identification was analyzed using the `glmer` function in the `lme4` R package. 
Estimates and odds ratio comparisons between the log and linear scales were calculated using the `emmeans` R package.
Each lineup plot evaluated was assigned a value based on the participant response (correct = 1, not correct = 0).
Define $Y_{ijkl}$ to be the event that participant $l = 1,...,N_{participant}$ correctly identifies the target plot for data set $k = 1,2$ with curvature combination $j = 1,2,3,4,5,6$ plotted on scale $i = 1,2$.
The binary response was analyzed using a generalized linear mixed model following a binomial distribution with a logit link function with a row-column blocking design accounting for the variation due to participant and data set respectively as 
\begin{equation}
\text{logit }P(Y_{ijk}) = \eta + \delta_i + \gamma_j + \delta \gamma_{ij} + s_l + d_k
\end{equation}
\noindent where

+ $\eta$ is the baseline average probability of selecting the target plot
+ $\delta_i$ is the effect of scale $i = 1,2$
+ $\gamma_j$ is the effect of curvature combination $j = 1,2,3,4,5,6$
+ $\delta\gamma_{ij}$ is the two-way interaction between the $i^{th}$ scale and $j^{th}$ curvature combination
+ $s_l \sim N(0,\sigma^2_\text{participant})$is the random effect for participant characteristics
+ $d_k \sim N(0,\sigma^2_{\text{data}})$ is the random effect for data specific characteristics. 

\noindent We assume that random effects for data set and participant are independent.

```{r}
glmm_mod <- glmer(binary_response ~ curvature*scale + 
                            (1 | participant_id) +
                            (1 | dataset_id),
                  data = lineup_model_data,
                  family = binomial(link = "logit"))

lineup_anova <- car::Anova(glmm_mod) |> 
  broom::tidy()
# summary(glmm_mod)

lineup_anova |> 
  knitr::kable(digits = 3, caption = "Lineups GLMM ANOVA")
```

```{r, eval = F, include = F, fig.align = 'center', fig.width = 12, fig.height = 6}
lineup_lsmeans <- emmeans(glmm_mod, specs = ~ "curvature:scale", type = "response") |> 
  as_tibble()

lineup_lsmeans |> 
  mutate(curvature = factor(curvature, levels = c("t-E_n-H", "t-H_n-E", "t-E_n-M", "t-M_n-E", "t-M_n-H", "t-H_n-M"))) %>%
  ggplot(aes(x = curvature, y = prob, group = scale)) +
  geom_bar(stat = "identity", fill = "lightgray") +
  # geom_text(aes(y = asymp.UCL + 0.05)) +
  facet_wrap(~scale, ncol = 1) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.1) +
  theme_bw() +
  scale_y_continuous("Probability of target panel detected", limit = c(-0.05,1.1), breaks = seq(0,1,0.2)) +
  scale_x_discrete("Curvature")
```


```{r, fig.align = 'center', fig.width = 12, fig.height = 6}
lineups_or <- emmeans(glmm_mod, ~ scale | curvature, type = "response") |> 
  pairs(reverse = T, infer = c(T,T)) |> 
  as_tibble()

# write.csv(lineups_or, here("analyses/01-lineups/results/lineups-odds-ratios.csv"), row.names = F, na =)

dodge <- position_dodge(width=0.9)
lineups_or_plot <- lineups_or %>%
  separate(curvature, into = c(NA, "target", NA, "null")) %>%
  mutate(null = factor(null, levels = c("E", "M", "H"), labels = c("High Curvature", "Medium Curvature", "Low Curvature")),
         target = factor(target, levels = c("E", "M", "H"), labels = c("High Curvature", "Medium Curvature", "Low Curvature"))) %>%
  ggplot(aes(x = odds.ratio, y = null, color = target, shape = target)) + 
  geom_point(position = dodge, size = 3) + 
  geom_errorbar(aes(xmin = asymp.LCL, xmax = asymp.UCL), position = dodge, width = .1) +
  geom_vline(xintercept = 1) +
  theme_bw()  +
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        legend.key.size = unit(1, "line"),
        legend.position = "bottom"
  ) +
  scale_y_discrete("Null Panel Difficulty", position = "left") +
  scale_x_continuous("Odds ratio (on log scale) \n (Log vs Linear)", trans = "log10", labels = scales::comma) + 
  scale_color_manual("Target Panel Difficulty", values = c("#004400", "#116611", "#55aa55")) + 
  scale_shape_discrete("Target Panel Difficulty")

lineups_or_plot
```

# Confidence Level Analysis

```{r}
#| fig-width: 12
#| fig-height: 8
# labellers for plotting
target_level <- c(
  "H" = "Low Target Curvature",
  "M" = "Moderate Target Curvature",
  "E" = "High Target Curvature"
)

null_level <- c(
  "H" = "Low Null Curvature",
  "M" = "Moderate Null Curvature",
  "E" = "High Null Curvature"
)

scale_level <- c("linear" = "Linear Scale",
                 "log" = "Log Scale")

lineup_model_data |> 
  mutate(target = factor(target, levels = c("H", "M", "E")),
         null = factor(null, levels = c("H", "M", "E")),
         conf_level = factor(conf_level, levels = c("Very Uncertain", "Uncertain", "Neutral", "Certain", "Very Certain")),
         binary_response = factor(binary_response, levels = c(1,0), labels = c("Correct", "Incorrect"))
         ) |>
  ggplot() +
  geom_mosaic(aes(x = product(binary_response),
             fill = conf_level),
             show.legend = T,
             color = "black") +
  facet_nested(null ~ scale + target,
               # labeller = label_both
             labeller = labeller(scale = scale_level,
                                 null = null_level,
                                 target = target_level)
             ) +
  scale_fill_brewer(palette = "PuOr",
                    guide = guide_legend(reverse = TRUE)) +
  labs(y = "",
       x = "Evaluation",
       fill = "How certain are you?") +
  theme_test() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

```{r}
confidence_data <- lineup_model_data |> 
  select(participant_id, curvature, target, null, scale, binary_response, conf_level, conf_level_num) |> 
  mutate(binary_response = factor(binary_response, levels = c(1,0), labels = c("Correct", "Incorrect"))) |> 
  mutate(conf_level_num2 = 6 - conf_level_num)

# linear model
library(lme4)
confidence_model <- lmer(conf_level_num2 ~ curvature*scale*binary_response + (1|participant_id),
                       data = confidence_data)
library(lmerTest)
anova(confidence_model)
plot(confidence_model)

library(emmeans)
emmeans(confidence_model, specs =~ binary_response|curvature) |> pairs()
emmeans(confidence_model, specs =~ scale|curvature) |> pairs()

emmeans(confidence_model, specs =~ binary_response) |> pairs()
emmeans(confidence_model, specs =~ scale) |> pairs()
```

# Rorschach lineups

```{r}
lineup_rorschach_data <- read_csv("lineup-rorschach-data.csv") |> 
  mutate(conf_level = factor(conf_level, levels = c("Very Certain", "Certain", "Neutral", "Uncertain", "Very Uncertain"))) |> 
  mutate(conf_level_num = as.numeric(conf_level))
lineup_rorschach_data
```

```{r}
# Calculate alphas
npanels = 20

calcAlpha <- function(data, m0 = 20) {
  
  K = nrow(data)
  c = m0/K
  zc = data |>
    select(response_no) |> 
    count(response_no) |> 
    arrange(-n) |> 
    filter(n >= c) |> 
    nrow()

  results <- estimate_alpha_numeric(Zc = zc,
                         c = c,
                         m0 = m0,
                         K  = K
                         )

  return(results$alpha)
}

alpha_values <- lineup_rorschach_data |> 
  nest(-target, -set, -scale) |> 
  arrange(target, set, scale) |> 
  mutate(alpha = map(.x = data,
                     .f = calcAlpha)
         ) |> 
  unnest(alpha) |> 
  mutate(target = fct_relevel(target,
                              c("H", "M", "E"))) |> 
  select(set, target, scale, alpha)

# unique(lineup_rorschach_data$dataset_id) |> length()
target_level <- c(
  "H" = "Low Curvature",
  "M" = "Medium Curvature",
  "E" = "High Curvature"
)

library(RColorBrewer)
nb.cols <- 16
mycolors <- colorRampPalette(brewer.pal(16, "Greens"))(nb.cols)

freqNum = function(data){
  data2 <- data |> 
    mutate(freq_order = 1:n())
  return(data2)
}

rorschach_summary <- lineup_rorschach_data |> 
  mutate(target = fct_relevel(target,
                              c("H", "M", "E"))) |> 
  group_by(evaluation_id, dataset_id, set, target, scale) |>
  mutate(weight = 1/n()) |>
  ungroup() |> 
  group_by(dataset_id, set, target, scale) |>
  count(response_no, wt = weight) |> 
  arrange(dataset_id, set, target, scale, -n) |> 
  nest(data = c(response_no, n)) |>
  mutate(data2 = map(data, freqNum)) |> 
  unnest(data2) |> 
  select(-data)

rorschach_summary |> 
  ggplot(aes(x = as.factor(set),
             y = n
             )
         ) +
  geom_bar(stat = "identity",
           position = "fill",
           show.legend = F,
           aes(fill = fct_rev(as.factor(freq_order)))
           ) +
  geom_text(data = alpha_values,
            aes(label = paste0("alpha ==", round(alpha, 2)), 
                y = 1.05
                ),
            size = 3,
            parse = TRUE
            ) +
  facet_grid(scale ~ target, 
             scales = "free_x",
             labeller = labeller(target = target_level)
             ) +
  theme_bw() +
  scale_fill_manual(values = mycolors) +
  labs(x = "Set",
       y = "Proportion of Selections",
       fill = "Identification")
```

# Explore Rorschach further

```{r}
names(lineup_rorschach_data)
names(lineup_model_data)


top_rorschach_avg <- rorschach_summary |> 
  group_by(dataset_id, set, target, scale) |> 
  mutate(pct_selected = n/sum(n)) |> 
  filter(freq_order == 1) |> 
  mutate(target = factor(target, 
                         levels = c("H", "M", "E"),
                         labels = c("Null: Low Curvature",
                                    "Null: Medium Curvature",
                                    "Null: High Curvature"
                                    )
                         )
  ) |> 
  rename(null = target) |> 
  arrange(scale, null, set) |> 
  group_by(null, scale) |> 
  summarize(avg_pct_selected = weighted.mean(x = pct_selected, w = n))




target_level <- c(
  "H" = "Low Curvature",
  "M" = "Medium Curvature",
  "E" = "High Curvature"
)

null_level <- c(
  "H" = "Low Curvature",
  "M" = "Medium Curvature",
  "E" = "High Curvature"
)
```


```{r}

accuracy_plot_linear <- lineup_model_data |> 
  filter(scale == "linear") |> 
  mutate(target = factor(target, 
                         levels = c("H", "M", "E"),
                         labels = c("Target: Low",
                                    "Target: Medium",
                                    "Target: High"
                                    )
                         ),
         null = factor(null, 
                       levels = c("H", "M", "E"),
                       labels = c("Null: Low Curvature",
                                  "Null: Medium Curvature",
                                  "Null: High Curvature"
                                  )
                       ),
         curvature_diff = case_when(curvature %in% c("t-E_n-H", "t-H_n-E") ~ "A - Large Diff",
                                    curvature %in% c("t-M_n-H", "t-H_n-M") ~ "B - Small Diff",
                                    curvature %in% c("t-E_n-M", "t-M_n-E") ~ "C - Small Diff"),
         curvature = factor(curvature, levels = curvature_order, 
                            labels = c("Target: High\nNull: Low",
                              "Target: Low \nNull: High",
                              "Target: Medium \nNull: Low",
                              "Target: Low \nNull: Medium",
                              "Target: High \nNull: Medium",
                              "Target: Medium \nNull: High")
                            )
         ) |> 
  ggplot(aes(x = target,
             fill = as.factor(binary_response))) +
  geom_bar(stat = "count",
           position = "fill") +
  # add rorschach
  geom_hline(data = top_rorschach_avg |> filter(scale == "linear"),
             aes(yintercept =  avg_pct_selected),
             color = "black",
             linetype = "dashed"
             ) +
  scale_fill_manual(values = c("darkgray", "#008b00"), label = c("Incorrect", "Correct")) +
  scale_y_continuous(label = scales::percent) +
  scale_x_discrete(position = "top") +
  facet_grid(scale ~ null,
             labeller = labeller(scale = scale_level),
             scales = "free_x"
             ) +
  # facet_grid(scale ~ curvature_diff,
  #            labeller = labeller(scale = scale_level,
  #                                curvature_diff = curvature_diff_level),
  #            scales = "free_x"
  #            ) +
  labs(y = "",
       x = "",
       fill = "Target Panel \nIdentification",
       title = str_c("Probability of Participants who ", "<span style = 'color:#008b00'>**correctly**</span>", " and ", 
       "<span style = 'color:#808080'>**incorrectly**</span>", " identified the target panel")) +
  theme_test() +
  theme(axis.ticks.x = element_blank(),
        plot.title = ggtext::element_markdown(size = 14),
        legend.position = "none",
        aspect.ratio = 1.5)

accuracy_plot_linear

ggsave(plot = accuracy_plot_linear, filename = "../images/1-linear.svg", dpi = 300, width = 8, height = 5)


accuracy_plot_log <- lineup_model_data |> 
  filter(scale == "log") |> 
  mutate(target = factor(target, 
                         levels = c("H", "M", "E"),
                         labels = c("Target: Low",
                                    "Target: Medium",
                                    "Target: High"
                                    )
                         ),
         null = factor(null, 
                       levels = c("H", "M", "E"),
                       labels = c("Null: Low Curvature",
                                  "Null: Medium Curvature",
                                  "Null: High Curvature"
                                  )
                       ),
         curvature_diff = case_when(curvature %in% c("t-E_n-H", "t-H_n-E") ~ "A - Large Diff",
                                    curvature %in% c("t-M_n-H", "t-H_n-M") ~ "B - Small Diff",
                                    curvature %in% c("t-E_n-M", "t-M_n-E") ~ "C - Small Diff"),
         curvature = factor(curvature, levels = curvature_order, 
                            labels = c("Target: High\nNull: Low",
                              "Target: Low \nNull: High",
                              "Target: Medium \nNull: Low",
                              "Target: Low \nNull: Medium",
                              "Target: High \nNull: Medium",
                              "Target: Medium \nNull: High")
                            )
         ) |> 
  ggplot(aes(x = target,
             fill = as.factor(binary_response))) +
  geom_bar(stat = "count",
           position = "fill") +
  # add rorschach
  geom_hline(data = top_rorschach_avg |> filter(scale == "log"),
             aes(yintercept =  avg_pct_selected),
             color = "black",
             linetype = "dashed"
             ) +
  scale_fill_manual(values = c("darkgray", "#008b00"), label = c("Incorrect", "Correct")) +
  scale_y_continuous(label = scales::percent) +
  scale_x_discrete(position = "top") +
  facet_grid(scale ~ null,
             labeller = labeller(scale = scale_level),
             scales = "free_x"
             ) +
  # facet_grid(scale ~ curvature_diff,
  #            labeller = labeller(scale = scale_level,
  #                                curvature_diff = curvature_diff_level),
  #            scales = "free_x"
  #            ) +
  labs(y = "",
       x = "",
       fill = "Target Panel \nIdentification") +
  theme_test() +
  theme(axis.ticks.x = element_blank(),
        plot.title = ggtext::element_markdown(size = 14),
        legend.position = "none",
        aspect.ratio = 1.5)

accuracy_plot_log

ggsave(plot = accuracy_plot_log, filename = "../images/1-log.svg", dpi = 300, width = 8, height = 5)
```






```{r}
# for reviewer

top_rorschach <- rorschach_summary |> 
  group_by(dataset_id, set, target, scale) |> 
  mutate(pct_selected = n/sum(n)) |> 
  filter(freq_order == 1) |> 
  mutate(target = factor(target, 
                         levels = c("H", "M", "E"),
                         labels = c("Null: Low Curvature",
                                    "Null: Medium Curvature",
                                    "Null: High Curvature"
                                    )
                         )
  ) |> 
  rename(null = target)

accuracy_plot_linear <- lineup_model_data |> 
  filter(scale == "linear") |> 
  mutate(target = factor(target, levels = c("H", "M", "E")),
         null = factor(null, levels = c("H", "M", "E")),
         curvature_diff = case_when(curvature %in% c("t-E_n-H", "t-H_n-E") ~ "A - Large Diff",
                                    curvature %in% c("t-M_n-H", "t-H_n-M") ~ "B - Small Diff",
                                    curvature %in% c("t-E_n-M", "t-M_n-E") ~ "C - Small Diff"),
         curvature = factor(curvature, levels = curvature_order, 
                            labels = c("Target: High\nNull: Low",
                              "Target: Low \nNull: High",
                              "Target: Medium \nNull: Low",
                              "Target: Low \nNull: Medium",
                              "Target: High \nNull: Medium",
                              "Target: Medium \nNull: High")
                            )
         ) |> 
  ggplot() +
  geom_bar(aes(x = set,
             fill = as.factor(binary_response)),
           stat = "count",
           position = "fill") +
  
  # add rorschach
  geom_hline(data = top_rorschach |> filter(scale == "linear"),
             aes(yintercept =  pct_selected),
             color = "orange3",
             linetype = "dashed"
             ) +
  
  #
  
  scale_fill_manual(values = c("darkgray", "#008b00"), label = c("Incorrect", "Correct")) +
  scale_y_continuous(label = scales::percent) +
  scale_x_discrete()  +
  facet_grid(target ~ null,
             labeller = labeller(target = target_level,
                                 null = null_level,
                                 .default = label_both
                                 ),
             scales = "free_x"
             ) +
  # facet_grid(scale ~ curvature_diff,
  #            labeller = labeller(scale = scale_level,
  #                                curvature_diff = curvature_diff_level),
  #            scales = "free_x"
  #            ) +
  labs(y = "",
       x = "Set",
       fill = "Target Panel \nIdentification",
       title = str_c("Probability of Participants who ", "<span style = 'color:#008b00'>**correctly**</span>", " and ", 
       "<span style = 'color:#808080'>**incorrectly**</span>", " identified the target panel"),
       subtitle = "Linear Scale") +
  theme_test() +
  theme(axis.ticks.x = element_blank(),
        plot.title = ggtext::element_markdown(size = 14),
        legend.position = "none",
        aspect.ratio = 1)

accuracy_plot_linear


accuracy_plot_log <- lineup_model_data |> 
  filter(scale == "log") |> 
  mutate(target = factor(target, levels = c("H", "M", "E")),
         null = factor(null, levels = c("H", "M", "E")),
         curvature_diff = case_when(curvature %in% c("t-E_n-H", "t-H_n-E") ~ "A - Large Diff",
                                    curvature %in% c("t-M_n-H", "t-H_n-M") ~ "B - Small Diff",
                                    curvature %in% c("t-E_n-M", "t-M_n-E") ~ "C - Small Diff"),
         curvature = factor(curvature, levels = curvature_order, 
                            labels = c("Target: High\nNull: Low",
                              "Target: Low \nNull: High",
                              "Target: Medium \nNull: Low",
                              "Target: Low \nNull: Medium",
                              "Target: High \nNull: Medium",
                              "Target: Medium \nNull: High")
                            )
         ) |> 
  ggplot() +
  geom_bar(aes(x = set,
             fill = as.factor(binary_response)),
           stat = "count",
           position = "fill") +
  
  # add rorschach
  geom_hline(data = top_rorschach |> filter(scale == "log"),
             aes(yintercept =  pct_selected),
             color = "orange3",
             linetype = "dashed"
             ) +
  scale_fill_manual(values = c("darkgray", "#008b00"), label = c("Incorrect", "Correct")) +
  scale_y_continuous(label = scales::percent) +
  scale_x_discrete(labels = c("Target: High\nNull: Low",
                              "Target: Low \nNull: High",
                              "Target: Medium \nNull: Low",
                              "Target: Low \nNull: Medium",
                              "Target: High \nNull: Medium",
                              "Target: Medium \nNull: High")) +
  labs(y = "",
       x = "",
       fill = "Target Panel \nIdentification") +
  theme_test() +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none",
        aspect.ratio = 1.5)

accuracy_plot_log


# unique lineups
lineup_model_data |> 
  select(set, null, target, dataset_id, correct) |> 
  filter(null == "M", target == "E") |> 
  distinct()

poo <- lineup_model_data |> 
  filter(null == "M", target == "E") |>
  filter(binary_response == 1) |> 
  separate_longer_delim(choice_reason, delim = ",") |> 
  count(set, scale, binary_response, choice_reason) |> 
  group_by(set, scale) |> 
  mutate(pct = n/sum(n)) |> 
  arrange(scale, set, -n)


# unique lineups
lineup_model_data |> 
  filter(null == "M", target == "E", scale == "linear") |> 
  count(set, response_no)
  


lineup_model_data |> 
  select(set, null, target, dataset_id, correct) |> 
  filter(null == "H", target == "M") |> 
  distinct()

poo <- lineup_model_data |> 
  filter(null == "H", target == "M") |>
  filter(binary_response == 1) |> 
  separate_longer_delim(choice_reason, delim = ",") |> 
  count(set, scale, binary_response, choice_reason) |> 
  group_by(set, scale) |> 
  mutate(pct = n/sum(n)) |> 
  arrange(scale, set, -n)
```




```{r}
# Overlay the counts per panel....

```

